---
title: "Hemisphere"
author: "WHMackenzie"
date: "2025-02-26"
output: html_document
---

```{r packages, include=FALSE}
library(magick)
library(hemispheR)
library(dplyr)
library("exifr")
require(tictoc)
require(data.table)
library(parallel)
library(foreach)
library(doParallel)  
library(tidyverse) # For data manipulation
library(imager) # For image display    
library(exifr) # For extracting metadata    
# ImageMagick is a command line tool for image manipulation. We will call ImageMagick from within R using the magick package, but first you'll need to download [https://imagemagick.org/script/download.php] and install ImageMagick on your machine.
    library(magick) # For image manipulation
    # Check to ensure that ImageMagick is installed.
    magick_config()$version 
     # For binarizing and calculating some canopy metrics, we will use Chiannuci's hemispheR package, which we need to install from the development version.
    library(devtools)
    #devtools::install_git("https://gitlab.com/fchianucci/hemispheR")
    library(hemispheR) # For binarization and estimating canopy measures
```

set up directories and read the mask
```{r setup, include=FALSE}
### Run the processing loop
    
    # Define the input path to the directory with panos
    #drive = "C:/Users/myarham/OneDrive - Government of BC/"
    drive = "D:/OneDrive - Government of BC/OffSite-Trials/Michelle/Lw_Understory/Insta360_Canopy_photos/"
    focal_path <- paste0(drive, "test folder") 
    list_of_panos <- list.files(focal_path) # Get the list of all panos
    length(list_of_panos)
    
    # Define the output directory path
    output <- paste0(drive, "Test Results") # Instantiate an empty object to receive the results

mask_path <- paste0(drive, "HemiPhotoMask.svg")

# Read the mask once
image_mask <- image_read(mask_path) %>%
  image_transparent("white") %>%
  image_convert("png")

# Create output directory if it doesn't exist
output_dir <- paste0(drive, "masked_hemispheres/")
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# Create intermediate results directory if it doesn't exist
intermediate_dir <- "./intermediate_results/"
if (!dir.exists(intermediate_dir)) {
  dir.create(intermediate_dir)
}
```

Loop through all images in test folder
Tried to parallelize the loop but was not successful due to magick package limitations.
Instead produce individual output files in loop and then combine into a single combined file after the loop.
This is more useful for adding new photos to existing processed images. 
```{r process image loop}
tic()
for (i in seq_along(list_of_panos)) {
  T1 <- Sys.time() # Used for time check

  # Construct the full path to the current image
  focal_image_path <- file.path(focal_path, list_of_panos[i])
  
  # Extract the base name (without extension) of the current image
  focal_image_name <- sub("\\.[^.]+$", "", basename(focal_image_path))
  
  # Print progress for debugging
  print(paste("Processing image:", focal_image_name))
  
  # Read EXIF data
  xmp_data <- read_exif(focal_image_path) %>%
    select(SourceFile, Make, Model, Megapixels, GPSLatitude, GPSLongitude)
  
  # Read and process the image
  pano <- image_read(focal_image_path)
  pano_width <- image_info(pano)$width
  image_heading <- 0 # Assuming heading is always 0
  
  pano_hemisphere <- pano %>%
    image_crop(geometry_size_percent(100, 50)) %>%
    image_resize(geometry_size_percent(100, 400)) %>%
    image_distort("Polar", c(0), bestfit = TRUE) %>%
    image_flip() %>%
    image_rotate(image_heading) %>%
    image_crop(paste0(pano_width, "x", pano_width, "-", pano_width/2, "-", pano_width/2))
  
  # Resize the mask to match pano width
  resized_mask <- image_resize(image_mask, geometry_size_pixels(width = pano_width, height = pano_width))
  
  masked_hemisphere <- image_mosaic(c(pano_hemisphere, resized_mask))
  
  masked_hemisphere_path <- file.path(output_dir, paste0(focal_image_name, "hemi_masked.jpg"))
  image_write(masked_hemisphere, masked_hemisphere_path)
  
  # Import and binarize the fisheye image
  fisheye <- import_fisheye(masked_hemisphere_path,
                            channel = '2BG',
                            circ.mask = list(xc = pano_width/2, yc = pano_width/2, rc = pano_width/2),
                            gamma = 2.2,
                            stretch = FALSE,
                            display = FALSE,
                            message = FALSE)
  
  binimage <- binarize_fisheye(fisheye,
                               method = 'Otsu',
                               zonal = FALSE,
                               manual = NULL,
                               display = FALSE,
                               export = TRUE)
  
  # Estimate canopy metrics
  gapfrac <- gapfrac_fisheye(binimage,
                             maxVZA = 90,
                             lens = "equidistant",
                             endVZA = 80,
                             display = FALSE,
                             message = FALSE)
  
  canopy_report <- canopy_fisheye(gapfrac)
  
  output_report <- xmp_data %>%
    bind_cols(canopy_report) %>%
    rename(GF = x, HemiFile = id)
  
  # Save intermediate results to disk
  intermediate_path <- file.path(intermediate_dir, paste0(focal_image_name, "_report.csv"))
  write.csv(output_report, intermediate_path, row.names = FALSE)
  
  # Calculate and print processing times
  T2 <- Sys.time()
  T_instance <- difftime(T2, T1, units = "secs")
  T_total <- difftime(T2, T0, units = "secs")
  T_average <- T_total / i
  
  print(paste0("Completed ", i, " of ", length(list_of_panos), " images in ", round(T_instance, 0), " seconds."))
  print(paste0("Estimated ", round(((length(list_of_panos) - i) * T_average) / 60, 1), " minutes remaining."))
}

# Combine intermediate results into a single CSV
intermediate_files <- list.files(intermediate_dir, full.names = TRUE)
output <- do.call(rbind, lapply(intermediate_files, read.csv))
toc()
# Write the final output to CSV
write.csv(output, "./canopy_output.csv", row.names = FALSE)
```

Loop through all images.
Tried to parallelize the loop but was not successful due to magick package limitations.
```{r}
# tic()
# # Initialize start time
# T0 <- Sys.time()
# # Initialize output data frame
# output <- data.frame()
# 
# for (i in seq_along(list_of_panos)) {
#   T1 <- Sys.time() # Used for time check
# 
#   # Construct the full path to the current image
#   focal_image_path <- file.path(focal_path, list_of_panos[i])
#   
#   # Extract the base name (without extension) of the current image
#   focal_image_name <- sub("\\.[^.]+$", "", basename(focal_image_path))
#   
#   # Print progress for debugging
#   print(paste("Processing image:", focal_image_name))
#   
#   # Read EXIF data
#   xmp_data <- read_exif(focal_image_path) %>%
#     select(SourceFile, Make, Model, Megapixels, GPSLatitude, GPSLongitude)
#   
#   # Read and process the image
#   pano <- image_read(focal_image_path)
#   pano_width <- image_info(pano)$width
#   image_heading <- 0 # Assuming heading is always 0
#   
#   pano_hemisphere <- pano %>%
#     image_crop(geometry_size_percent(100, 50)) %>%
#     image_resize(geometry_size_percent(100, 400)) %>%
#     image_distort("Polar", c(0), bestfit = TRUE) %>%
#     image_flip() %>%
#     image_rotate(image_heading) %>%
#     image_crop(paste0(pano_width, "x", pano_width, "-", pano_width/2, "-", pano_width/2))
#   
#   # Resize the mask to match pano width
#   resized_mask <- image_resize(image_mask, geometry_size_pixels(width = pano_width, height = pano_width))
#   
#   masked_hemisphere <- image_mosaic(c(pano_hemisphere, resized_mask))
#   
#   masked_hemisphere_path <- file.path(output_dir, paste0(focal_image_name, "hemi_masked.jpg"))
#   image_write(masked_hemisphere, masked_hemisphere_path)
#   
#   # Import and binarize the fisheye image
#   fisheye <- import_fisheye(masked_hemisphere_path,
#                             channel = '2BG',
#                             circ.mask = list(xc = pano_width/2, yc = pano_width/2, rc = pano_width/2),
#                             gamma = 2.2,
#                             stretch = FALSE,
#                             display = FALSE,
#                             message = FALSE)
#   
#   binimage <- binarize_fisheye(fisheye,
#                                method = 'Otsu',
#                                zonal = FALSE,
#                                manual = NULL,
#                                display = FALSE,
#                                export = TRUE)
#   
#   # Estimate canopy metrics
#   gapfrac <- gapfrac_fisheye(binimage,
#                              maxVZA = 90,
#                              lens = "equidistant",
#                              endVZA = 70, # 70 is default value
#                              display = FALSE,
#                              message = FALSE)
#   
#   canopy_report <- canopy_fisheye(gapfrac)
#   
#   output_report <- xmp_data %>%
#     bind_cols(canopy_report) %>%
#     rename(GF = x, HemiFile = id)
# 
#   # Save intermediate results to disk
#   intermediate_path <- file.path(output_dir, paste0(focal_image_name, "_report.csv"))
#   write.csv(output_report, intermediate_path, row.names = FALSE)
# 
#   # Write the final output to CSV
# #write.csv(output, "./canopy_output.csv", row.names = FALSE)
#   output <- bind_rows(output, output_report)
#   
#   # Calculate and print processing times
#   T2 <- Sys.time()
#   T_instance <- difftime(T2, T1, units = "secs")
#   T_total <- difftime(T2, T0, units = "secs")
#   T_average <- T_total / i
#   
#   print(paste0("Completed ", i, " of ", length(list_of_panos), " images in ", round(T_instance, 0), " seconds."))
#   print(paste0("Estimated ", round(((length(list_of_panos) - i) * T_average) / 60, 1), " minutes remaining."))
#   gc()
# }
#  
# fwrite(output, paste0(drive,"./canopy_output_WHM.csv")) 
# # Read intermediate results and combine them
#   output2 <- do.call(rbind, lapply(output, read.csv))
# # Write the final output to CSV
# fwrite(output2, paste0(drive,"./canopy_output_WHM.csv")) 
# toc()
```

